---
title: ""
output:
  html_document
theme: cosmo
---

## R `apply` family {.tabset}

We will use a subset of the `mtcars` dataset

```{r}
data <- mtcars[, c(1,2,10)]
```

### apply()

Apply works on arrays and matrices <br>

`apply(x, MARGIN, fun, ...)`

* `x`: array or matrix
* `MARGIN`: When `=1`, applies over rows. `=2` applies over columns. `=c(1,2)` applies over rows and columns 
* `fun`: the function to be applied

Example: <br>
```{r}
# Creating a matrix
matrice <- as.matrix(data)

# Mean of each column
apply(matrice, MARGIN = 2, mean)

```

### lapply()

Lapply works on lists, dataframes and vectors, returning a list of the same length as x <br>

`lapply(x, fun, ...)`

```{r}
# sum of each column
lapply(data, sum)

# creating 1 dataframe for each gear
three <- subset(data, gear == 3)
four <- subset(data, gear == 4)
five <- subset(data, gear == 5)

# sum first column of each dataframe
data.all <- list(three, four, five)    # create a list with all 3 dataframes
first.col <- lapply(data.all,"[", 1)   # create a list with the first column of each dataframe
lapply(first.col, sum)

```



### sapply()

Sapply is a simplified version of `lapply()`, because instead of returning a list, returns a vector (or an array if `simplify = array`, or a list just like `lapply()` if `simplify = F`) <br>

`sapply(x, fun, ..., simplify = T)` <br>

```{r}
# return the first element of the second row on each dataframe - data.all
sapply(data.all, "[", 2, 1)

# the same as above but as a list
sapply(data.all, "[", 2, 1, simplify = F)

```



### mapply()

"Multivariate apply".




















